Community-conditioned language models for linguistic variation

### 2021-11-29 paper/paper1.tex
lstm vs transformer

We experiment with two kinds of model architecture: a simple
unidirectional LSTM \citep{Hochreiter1997} and a Transformer
\citep{Vaswani2017}.  In either case, the model is organised as a
standard $3$-layer neural sequence encoder, where the input for the
$t$th timestep of the $n+1$st layer is the $t$th hidden state of the 
$n$th layer.
As usual, the input to the first layer, is a sequence of tokens,
encoded with a trainable embedding layer over a pre-determined vocabulary.
At the other end, word tokens are predicted using a softmax projection 
layer. What we have described so far does not take community into account
and as such we call them \emph{uncodnitioned models}, but the same
encoder architecture also forms the core of our conditioned models.

@@

>>>>

We experiment with two kinds of model architecture: a simple
unidirectional LSTM \citep{Hochreiter1997} and a Transformer
\citep{Vaswani2017}. Even though the transformer is dominating LSTM in
the state of the art, it achieves its dominance partly thanks to the
availability of very large data sets \jp{Bill: can you add relevant citations?
  (BERT, GPT3, etc.)},
which are not available to us.\footnote{Fine-tuning existing models
  falls outside our methodology, which consists in fundamentally
  changing the structure of the network by concatenating
  community embeddings.}.

In either case, the model is organised as a
standard $3$-layer neural sequence encoder, where the input for the
$t$th timestep of the $n+1$st layer is the $t$th hidden state of the 
$n$th layer.
As usual, the input to the first layer, is a sequence of tokens,
encoded with a trainable embedding layer over a pre-determined vocabulary.
At the other end, word tokens are predicted using a softmax projection 
layer. What we have described so far does not take community into account
and as such we call them \emph{uncodnitioned models}, but the same
encoder architecture also forms the core of our conditioned models.

@@

We additionally observe that the best LSTM model achieves better
performance than the best Transformer model. This partly contradicts
the fact that Transformer-based models are the ones used in
state-of-the art models for mainstream tasks. We propose that the
cause is that we are training all models from scratch on a relatively
small dataset. In this situation, the relatively smaller number of
parameters of the LSTM means that it can generalise better than the
Transformer.
